---
title: "Plot Manuscript Figures -- Salmonid Culverts"
author: "Kelly"
date: "2022-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)

####
#read in time series model fit
stanMod <- readRDS(here("Scripts","timeseries_model","modelFit_20221118_notflowcorrected_greater1percent.RDS"))

f <- stanMod[3][[1]]  #data in
s <- stanMod[[2]] #fitted model itself
```

Summarize expected values at each time/station/species/creek by capturing parameter estimates for mean and standard deviation from model fit.

```{r, echo = F, warning = F, message= F}

resOut <- expand_grid(time_idx = 1:length(unique(f$time_idx)),
                      station_idx = 1:length(unique(f$station_idx)),
                      species_idx = 1:length(unique(f$species_idx)),
                      creek_idx = 1:length(unique(f$creek_idx))) %>%  
  mutate(mean_est = rstan::summary(s, par = "mu")$summary[,1],
         ci25 = rstan::summary(s, par = "mu")$summary[,5],
         ci75 = rstan::summary(s, par = "mu")$summary[,7]) %>% 
  left_join(f %>% dplyr::select(creek, species, creek_idx, species_idx) %>% distinct()) %>% 
  right_join(f %>% dplyr::select(meandnaconc, station_idx, time_idx, creek_idx, species_idx)) %>% 
  left_join(data.frame(sigma_dna = rstan::summary(s, par = "sigma_dna")$summary[,1], 
                       species_idx = 1:1:length(unique(f$species_idx))))
```

Sample posterior for expected value +/- observation variance at each time/station/species/creek

```{r, echo = F, warning = F, message = F}

  #function to generate samples drawn from normal distribution,
  #the params of which are the posterior distribs for mean and SD
  #this captures both process and observation variance
  getPostPred <- function(modname, time_idx, station_idx, species_idx, creek_idx){
        s = modname
        n = length(unlist(rstan::extract(s, par = "mu[1,1,1,1]")))
        f_mu = paste0("mu[",
                      time_idx,",",
                      station_idx,",",
                      species_idx,",",
                      creek_idx,"]")
        f_s = paste0("sigma_dna[",species_idx,"]")
        
        postsamples <- rnorm(n, 
                             unlist(rstan::extract(s, par = f_mu)),
                             unlist(rstan::extract(s, par = f_s))
        ) 
        
        return(quantile(postsamples, c(.025, .05, .25, .5, .75, .9, .975)))
  }

  #summarize and save in this object
  ppOut <- expand_grid(time_idx = 1:length(unique(f$time_idx)),
                         station_idx = 1:length(unique(f$station_idx)),
                         species_idx = 1:length(unique(f$species_idx)),
                         creek_idx = 1:length(unique(f$creek_idx)))
    ppOut[,5:11] <- 0.0
    colnames(ppOut)[5:11] <- paste0("pp_", c("025", "05", 25, 50, 75, 90, 975))
    
    for (i in 1:nrow(ppOut)){
      ppOut[i,5:11] <- as.list(getPostPred(s,
                                           ppOut$time_idx[i], 
                                           ppOut$station_idx[i], 
                                           ppOut$species_idx[i], 
                                           ppOut$creek_idx[i]))
    }
```

Make combined-species plot of trends over time

```{r, echo = F, warning = F, message = F}
  multispeciesTrendsplot <- f %>%
    mutate(logY = log(meandnaconc)) %>%
    right_join(resOut) %>% # get individual expected values for each sample
    # mutate(month = ifelse(time_idx < 11, time_idx + 2, time_idx - 10),
    #        # month = as.factor(month)
    # ) %>% 
    left_join(ppOut) %>% # get posterior estimates from full model (process + obs variance)
    drop_na() %>% 
    mutate(station = ifelse(station_idx == 1, "Downstream", "Upstream")) %>%
    mutate(facetorder = factor(creek, levels=c('Padden','Portage','Chuckanut','Squalicum'))) 

pad_only <- multispeciesTrendsplot %>% filter(creek=="Padden")

multispeciesTrendsplot <- ggplot() +  #treating meandnaconc as observations
    geom_rect(data=pad_only, aes(xmin = 8, xmax = 12, ymin = -Inf, ymax = Inf), alpha = .01, color=NA, fill = "grey") +
    geom_point(data=multispeciesTrendsplot, aes(x = newtime, y = log(meandnaconc), color = station), alpha = .2) +
    geom_point(data=multispeciesTrendsplot, aes(x = newtime, y = mean_est, color = station), size = 1.5, alpha = .5) + #mean estimates of mu
    geom_smooth(data=multispeciesTrendsplot, aes(x = newtime, y = mean_est, color = station), se = F, size = 1, alpha = .5, span = .2) +
    geom_segment(data=multispeciesTrendsplot, aes(x = newtime, xend = newtime, y = pp_025, yend = pp_975, color = station), size = .3,  alpha = .2) + #full error model creates CI
    geom_segment(data=multispeciesTrendsplot, aes(x = newtime, xend = newtime, y = pp_25, yend = pp_75, color = station), size = .7, alpha = .4) +
    facet_grid(species~facetorder) +
    xlab("Date (YY-MM)") + ylab("Log DNA Concentration (copies/L)") +
    labs(color='') +
    # ggtitle(paste0(unique(f$species)[i], "\n(predicted mean +/- 95 CI)")) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          legend.position = "bottom",
          legend.box.spacing = unit(0, "mm")) + 
    scale_x_discrete(guide = guide_axis(angle = -45))  
    
    #ggsave(multispeciesTrendsplot, file = here("Output/Figures/multispeciesTrends_notflowcorrected_greater1percent.png"), width=9, height=7)

```

Make species-specific plots and save as ggplot objects

```{r, echo = F, warning = F, message = F}
 
    for (i in 1:length(unique(f$species))){
      assign(paste0("p",i), value = 
               resOut %>%
               mutate(logY = log(meandnaconc)) %>%
               mutate(month = ifelse(time_idx < 11, time_idx + 2, time_idx - 10)) %>% 
               left_join(ppOut) %>% 
               mutate(station = ifelse(station_idx == 1, "Downstream", "Upstream")) %>%
               filter(species_idx == i) %>%
               ggplot(aes(x = month, y = log(meandnaconc), color = station)) +
               geom_point(alpha = .2) +
               geom_point(aes(x = month, y = mean_est, color = station), size = 1.5, alpha = .5) +
               #geom_smooth(aes(x = month, y = mean_est, color = station), se = F, size = 1, alpha = .5, span = .2) +
               geom_segment(aes(x = month, xend = month, y = pp_025, yend = pp_975, color = station), size = .3, alpha = .2) +
               geom_segment(aes(x = month, xend = month, y = pp_25, yend = pp_75, color = station), size = .7, alpha = .4) +
               facet_grid(~creek) +
               xlab("Month") + ylab("Log DNA Concentration (copies/L)") +
               ggtitle(paste0(unique(f$species)[i], "\n(predicted mean +/- 95 CI)")) +
               theme_bw() +
               theme(panel.grid.minor = element_blank(),
                     panel.grid.major.x = element_blank()) +
               scale_x_continuous(breaks = 1:12,
                                  labels = as.character(1:12))
             
      )
    }

# ggsave(p1, file = here("Output/Figures/Oclarkii_timeseries_notflowcorrected.png"))
# ggsave(p2, file = here("Output/Figures/Okisutch_timeseries_notflowcorrected.png"))
# ggsave(p3, file = here("Output/Figures/Omykiss_timeseries_notflowcorrected.png"))

```

Effect of construction itself on Padden Creek

```{r, echo = F, warning = F, message = F}
#recover posterior estimates of gamma term
    gamma_est <- expand_grid(time_idx = 2:(length(unique(f$time_idx))),
                               species_idx = 1:length(unique(f$species_idx)),
                               #creek_idx = 1:length(unique(f$creek)),
                               #station_idx = 1:length(unique(f$station_idx)),
                               construction_idx = 1:length(unique(f$construction_idx))
                               ) %>% 
      mutate(gamma_mean = NA,
             gamma_025 = NA,
             gamma_975 = NA)
    
    for (i in 1:nrow(gamma_est)){
      a <- paste0("gamma[", 
                  gamma_est$time_idx[i]-1, ",",
                  gamma_est$species_idx[i], ",",
                  #gamma_est$creek_idx[i], ",",
                  #gamma_est$station_idx[i], ",",
                  gamma_est$construction_idx[i], "]")
      
      gamma_est$gamma_mean[i] <- rstan::summary(s, par = a)$summary[,1]
      gamma_est$gamma_025[i] <- rstan::summary(s, par = a)$summary[,4]
      gamma_est$gamma_975[i] <- rstan::summary(s, par = a)$summary[,8]
    }

  # link to information about sampling dates  
  gamma_est <- f %>% select(newtime, time_idx) %>%
    distinct() %>%
    mutate(month = ifelse(time_idx < 11, time_idx + 2, time_idx-10)) %>% 
    left_join(gamma_est)

  Nspp <- length(unique(f$species))
  

# gamma_est %>% 
#   left_join(tibble(species_idx = 1:Nspp, jitter = (1:Nspp - median(1:Nspp))*0.07)) %>% 
#   left_join(f %>% dplyr::select(species_idx, species) %>% distinct()) %>% 
#   mutate(jitter_time = time_idx + (jitter)) %>% 
#   filter(gamma_mean != 0) %>% 
  # filter(construction_idx == 2, species_idx < 4 & month %in% c(1:12)) %>% 
  # ggplot(aes(x = jitter_time, y = gamma_mean, color = species)) +
  #   geom_point() +
  #   geom_segment(aes(x = jitter_time, xend = jitter_time, y = gamma_025, yend = gamma_975)) +
  #   scale_x_continuous(breaks = c(7:12),
  #                    labels = c(9:12,1,2)) +
  #   xlab("Month") + ylab("Gamma")

colorPal <- c("#8b62ca",
              "#55b04a",
              "#c760a3",
              "#9cb24e",
              "#688bcd",
              "#c7a03d",
              "#4bb092",
              "#6d732f")

#percent change due to gamma, as expressed by:
# gamma_mean/(mean_est-gamma_mean)
(construction_effect <- gamma_est %>% 
  filter(gamma_mean!= 0) %>% 
  left_join(tibble(species_idx = 1:Nspp, jitter = (1:Nspp - median(1:Nspp))*0.07)) %>% 
  left_join(f %>% dplyr::select(species_idx, species) %>% distinct()) %>% 
  mutate(jitter_time = time_idx + (jitter)) %>% 
  left_join(resOut %>% filter(creek_idx == 3, station_idx == 1) %>% dplyr::select(time_idx, species_idx, mean_est)) %>% distinct() %>% 
  mutate(gamma_mean_pct = gamma_mean/(mean_est-gamma_mean)*100,
         gamma_025_pct = gamma_025/(mean_est-gamma_mean)*100,
         gamma_975_pct = gamma_975/(mean_est-gamma_mean)*100) %>% 
  ggplot(aes(x = jitter_time, y = gamma_mean_pct, color = species)) +
  geom_point() +
  geom_segment(aes(x = jitter_time, xend = jitter_time, y = gamma_025_pct, yend = gamma_975_pct)) +
  scale_x_continuous(breaks = c(7:12),
                     labels = c(9:12,1,2)) +
  theme_bw() +
  labs( x="Month", y= "Percent Change due to Culvert Removal", title="Effect of Construction in Padden") +
  scale_color_discrete(type = colorPal))

#ggsave(construction_effect, file = here("Output/Figures/Construction_effect_notflowcorrected_greater1percent.png"))

```

Effect of culvert -- difference between eta terms above vs. below

```{r, echo = F, warning = F, message = F}


delta <- expand_grid(time_idx = 2:(length(unique(f$time_idx))),
                         #station_idx = 1:length(unique(f$station_idx)),
                         species_idx = 1:length(unique(f$species_idx)),
                         creek_idx = 1:length(unique(f$creek))
                   ) %>% 
  mutate(delta_mean = NA,
         delta_025 = NA,
         delta_975 = NA)

for (i in 1:nrow(delta)){
  a <- paste0("delta[", 
              delta$time_idx[i]-1, ",",
              delta$species_idx[i], ",",
              delta$creek_idx[i],"]")
  
  delta$delta_mean[i] <- rstan::summary(s, par = a)$summary[,1]
  delta$delta_025[i] <- rstan::summary(s, par = a)$summary[,4]
  delta$delta_975[i] <- rstan::summary(s, par = a)$summary[,8]
}

  # link to information about sampling dates  
delta <- f %>% select(newtime, time_idx) %>%
    distinct() %>%
    mutate(month = ifelse(time_idx < 11, time_idx + 2, time_idx-10)) %>% 
    left_join(delta) %>% 
    left_join(resOut %>% dplyr::select(time_idx, species_idx, creek_idx, station_idx, mean_est) %>% filter(station_idx == 1) %>% distinct()) %>% 
    mutate(delta_mean_pct = delta_mean/(mean_est-delta_mean)*100,  #scale to make percent change
           delta_025_pct = delta_025/(mean_est-delta_mean)*100,
           delta_975_pct = delta_975/(mean_est-delta_mean)*100) %>% 
    left_join(f %>% dplyr::select(species_idx, species, creek_idx, creek) %>% distinct())
  
  (culvert_effect <-  delta %>% 
      filter(mean_est > 2) %>% #Filter out very low-conc observations, because ratios make these misleading 
      ggplot(aes(x = as.factor(month), y = delta_mean_pct)) +
      geom_boxplot(fill = "skyblue") +
      xlab("Month") + ylab("Mean difference upstream - downstream \n(% of downstream concentration)") +
      theme_bw()
      )  #here, just taking means, and summarizing across species and creeks w boxplot
  
 # ggsave(culvert_effect, file = here("Output/Figures/culvert_effect_notflowcorrected.png"))

  (culvert_effect_creek_species <-  delta %>% 
      filter(mean_est > 2) %>% #Filter out very low-conc observations, because ratios make these misleading 
      drop_na() %>% 
      ggplot(aes(x = as.factor(month), y = delta_mean_pct)) +
      geom_point() +
      facet_grid(species~creek) +
      xlab("Month") + ylab("Mean difference upstream - downstream \n(% of downstream concentration)") +
      theme_bw()
  )
  
  # ggsave(culvert_effect_creek_species, file = here("Output/Figures/culvert_effect_creek_species_notflowcorrected.png"))
  
```

```{r make padden plot of just rainbow and add in hatchery releases}
# 
# pad_only <- multispeciesTrendsplot %>% filter(creek=="Padden")
# pad_mykiss <- pad_only %>% filter(species=="Oncorhynchus mykiss")
# 
# release_dates <- c()
# 
# ggplot(pad_mykiss) +  #treating meandnaconc as observations
#     geom_rect(aes(xmin = 8, xmax = 12, ymin = -Inf, ymax = Inf), alpha = .01, color=NA, fill = "grey") +
#     geom_point(aes(x = newtime, y = log(meandnaconc), color = station), alpha = .2) +
#     geom_point(aes(x = newtime, y = mean_est, color = station), size = 1.5, alpha = .5) + #mean estimates of mu
#     geom_smooth(aes(x = newtime, y = mean_est, color = station), se = F, size = 1, alpha = .5, span = .2) +
#     geom_segment(aes(x = newtime, xend = newtime, y = pp_025, yend = pp_975, color = station), size = .3,  alpha = .2) + #full error model creates CI
#     geom_segment(aes(x = newtime, xend = newtime, y = pp_25, yend = pp_75, color = station), size = .7, alpha = .4) +
#     xlab("Date (YY-MM)") + ylab("Log DNA Concentration (copies/L)") +
#     theme_bw() +
#     theme(panel.grid.minor = element_blank(),
#           panel.grid.major.x = element_blank(),
#           legend.position = "bottom",
#           legend.box.spacing = unit(0, "mm")) + 
#     scale_x_discrete(guide = guide_axis(angle = -45))

```
