---
title: "process_qpcr_data"
author: "Eily Allan"
date: "6/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
#library(MCMCpack) #for rdirichelet function
library(rstan)
#library(shinystan)
#library(bayesplot)
library(here)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

source(here("Scripts","qm_qpcr_model", "simFunctions.R")) 
```

Here, we are going to take the Ct values from the qpcr instrument and convert them into copy numbers. We will use the standard curve to do this. 

## Import data

We have a spreadsheet with our results of both standard curves and samples Ct values. We also have another spreadsheet with dilution information - we checked for inhibition before running the assay and if a sample was inhibited, we diluted until it was no longer inhibited (using IPC kit). So we will eventually need to multiply by the dilution factor after we move from Ct space to copy/uL space. 

```{r read files}

qpcr_meta <- read.csv(here("Input","qpcr","qpcr.dilutions.csv"))
qPCRdata <- read.csv(here("Input","qpcr","cutthroat_qpcr_results.csv"))

# join these together so the data has the dilution factor attached
qPCRdata <- qPCRdata %>% 
  left_join(qpcr_meta, by="Sample") 

```

Split data into standard curve and environmental samples. 

```{r}

# Standard curve data: remove standards not assigned a Ct value (undetermined)
# for now, only look at plate 1 because 2 and 3 are jacked up 

stdCurve <- qPCRdata %>% 
  filter(Type == "STANDARD" & Ct != "Undetermined" & Plate == 1) %>% 
  mutate(Ct = as.numeric(Ct),
         Sample = as.numeric(Sample))

# Environmental data: remove samples not assigned a Ct value (undetermined) 
# add column for technical replicate for each biological replicate

envir_qPCR <- qPCRdata %>% 
  filter(Type != "STANDARD") %>% 
  filter(Ct != "Undetermined") %>% 
  mutate(Ct = as.numeric(Ct),
         Final.qpcr.Dilution = as.numeric(Final.qpcr.Dilution)) %>% 
  arrange(Sample) %>% 
  separate(Sample, into = c("time", "site", "station", "biolRep"), remove = F) %>% 
  group_by(time, site, station, biolRep) %>% 
  mutate(tech_rep = 1:n())
```

Check std curves
```{r}
stdCurve %>% 
  ggplot(aes(x= log(Sample), y = Ct, color = as.factor(Plate))) +
    geom_point() +
    geom_smooth(method = "lm")

```

Because Ole and Ryan are in a math cult, we are not going to use a regular old linear model. Instead, we can have the variance change with the mean, which we want to capture because we expect detections to be more stochastic at low copy numbers. 


```{r qPCR calibrations setup}

stan_data <- list()
#add qPCR std curve
  stan_data$N_std_curve <- length(stdCurve$Ct)
  stan_data$N_std_samples <- length(unique(stdCurve$Sample))
  stan_data$std_sample_idx <- match(stdCurve$Sample, unique(stdCurve$Sample)) #note, assumes single array of dilutions
  stan_data$known_concentration <- unique(stdCurve$Sample)
  stan_data$y_ct_std_curve <- stdCurve$Ct

#add qPCR from envir samples
  stan_data$N_qPCR_envir <- nrow(envir_qPCR)
  stan_data$N_biol_estimates <- length(unique(envir_qPCR$Sample))
  stan_data$sp_idx_qPCR <- rep(1, stan_data$N_biol_estimates) #species idx; will need to change when we have multiple reference species
  stan_data$obs_samp_idx <- match(envir_qPCR$Sample, unique(envir_qPCR$Sample))
  stan_data$obs_samp_small_idx <- unique(envir_qPCR$Sample)
  stan_data$y_ct_envir <- envir_qPCR$Ct

```


qPCR calibration and estimating absolute concentrations from enviro samples using qPCR.

```{r fit qPCR calibration model and estimate concentration of enviro samples, results='hide'}

#TODO modify model to handle multiple different std curves

qMod = stan(file = here("Scripts", "qm_qpcr_model","qPCR_calibration.stan") ,data = stan_data,
                     verbose = FALSE, chains = 3, thin = 1,
                     warmup = 500, iter = 2000,
                     control = list(adapt_init_buffer = 175,
                                    max_treedepth=12,
                                    stepsize=0.01,
                                    adapt_delta=0.7,
                                    metric="diag_e"),
                     #pars = stan_pars,
                     refresh = 10,
                     boost_lib = NULL,
                     #init = stan_init_f2(n.chain=N_CHAIN,N_species=N_species),
                     sample_file = "temp/tmp.csv"
      )

#plot(qMod, par = c("beta_std_curve_1", "sigma_std_curve", "sigma_envir_qPCR"))
# launch_shinystan(qMod)
```

Visualize std curve with CI

```{r}
pp_beta0 <- extract(qMod, "beta_std_curve_0")$beta_std_curve_0
pp_beta1 <- extract(qMod, "beta_std_curve_1")$beta_std_curve_1
pp_sigma <- extract(qMod, "sigma_std_curve")$sigma_std_curve

conc <- runif(1000, 0, 5)
mu <- sample(pp_beta0, 1000) + (sample(pp_beta1, 1000)*conc)
y<-NA
for(i in 1:1000){y[i] <- rnorm(1, mean = mu[i], sd = pp_sigma[i])}

data.frame(y, conc) %>% 
  ggplot(aes(x = conc, y = y)) +
    geom_point(alpha = .5)

```

Map our actual environmental samples on.

```{r plot qPCR fit}
results_qPCR <- envir_qPCR %>%
  ungroup() %>% 
  filter(tech_rep == 1) %>%
  mutate(bayes_est = 10^(summary(qMod, par = "envir_concentration")$summary[,1]),
         bayes_25pct = 10^(summary(qMod, par = "envir_concentration")$summary[,5]),
         bayes_75pct = 10^(summary(qMod, par = "envir_concentration")$summary[,7]),
         copy_L_bayes_est = bayes_est*Final.qpcr.Dilution*100,
         copy_L_bayes_25pct = bayes_25pct*Final.qpcr.Dilution*100,
         copy_L_bayes_75pct = bayes_75pct*Final.qpcr.Dilution*100)
         

results_qPCR %>% 
  ggplot(aes(x = station, y = log(copy_L_bayes_est))) +
    geom_point() +
    geom_segment(aes(x = station, xend = station, y = log(copy_L_bayes_25pct), yend = log(copy_L_bayes_75pct))) +
    facet_wrap(~time) +
  labs(y= "Log Copies/L water", title = "Cutthroat Trout")

```




```{r environmental samples}

envir_qPCR <- qPCRdata %>% 
  filter(Type != "STANDARD") %>% 
  filter(Ct != "Undetermined") %>% 
  mutate(Ct = as.numeric(Ct)) %>% 
  arrange(Sample) %>% 
  separate(Sample, into = c("time", "site", "station", "biolRep"), remove = F) %>% 
  group_by(time, site, station, biolRep) %>% 
  mutate(tech_rep = 1:n())